Compiling refinement types to dependent types

### 2020-07-25 tex/paper.tex
More styling fixes
Less "we", more passive voice.

The syntax of the core language is presented in \figref{core_syntax},
the core $\lambda C$ typing rules are defined in \figref{core_typing}
and the rules for our extensions are defined in \figref{core_typing_exts}.
Note that the \textsc{CT-ADTCase} rule uses the equality $\equiv$,
which is a derived form defined later.

@@

Since $\Tranctx$ is defined on the derivations of $\nonterm G \ctxok$,
we consider the last rule used in a derivation:
\begin{itemize}
  \item \textsc{TCTX-Empty}.
    \begin{align*}
      & \inferrule{~}
                  {\emptyset \ctxok}
          \longmapsto
        \emptyset.
    \end{align*}
    The base case is trivial: we map the empty context to the empty context.
  \item \textsc{TCTX-Bind}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\ \nonterm G \vdash \nonterm{ts}}
                  {\nonterm G , \gterm x : \nonterm{ts} \ctxok}
          \longmapsto
        \Tranctx(\nonterm G \ctxok), x : \tranty(\nonterm G \vdash \nonterm{ts}).
    \end{align*}

@@

More formally, we consider the last rule used in the derivation of $\nonterm G \vdash \nonterm{ts}$:
\begin{itemize}
  \item \textsc{TWF-TrueRef}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok}
                  {\nonterm G \vdash \{ v : B | \top \}}
          \longmapsto
        \inferrule{\cdots}
                  {\nonterm{GC} \vdash (\hat B, \unitc \equiv \unitc) : \star}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok).
    \end{align*}

    We do not mention the premises of the translated rule.
    By \cref{lma:non_dep_pair_typing}
    $(\hat B, \unitc \equiv \unitc)$ has type $\star$ in $\nonterm{GC}$
    if $\hat B$ and $\unitc \equiv \unitc$ are well-typed in $\nonterm{GC}$,
    which, in turn, is easily checked.
    Thus, there exists a derivation of $\nonterm{GC} \vdash (\hat B, \unitc \equiv \unitc) : \star$,
    and it is effectively constructible~--- we just omit it for brevity.
  \item \textsc{TWF-Base}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\\\
                   \nonterm G ' \vdash \nonterm{es}_1 : \{ v_2 : B' | \nonterm{r}_1 \} \\\\
                   \nonterm G ' \vdash \nonterm{es}_2 : \{ v_2 : B' | \nonterm{r}_2 \} \\\\
                   \gterm x \fresh}
                  {\nonterm G \vdash \{ v : B | \nonterm{es}_1 = \nonterm{es}_2 \}}
          \longmapsto
        \inferrule{\Tranterm(\nonterm G ' \vdash \nonterm{es}_1 : \{ v_2 : B' | \nonterm{r}_1 \}) \\\\
                   \Tranterm(\nonterm G ' \vdash \nonterm{es}_2 : \{ v_2 : B' | \nonterm{r}_2 \}) \\\\
                   \gterm x \fresh \\\\
                   \cdots}
                  {\nonterm{GC} \vdash (v : \hat B \dast \lambda v : \hat B. \nonterm{ec}_1 \equiv \nonterm{ec}_2) : \star}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok), \\
                   &\nonterm G ' = \nonterm G , v : \{ v_1 : B | \top \}, \\
                   &\nonterm{ec}_i =
                     \tranterm(\nonterm G ' \vdash \nonterm{es}_i : \{ v_2 : B' | \nonterm{r}_i \}).
    \end{align*}

@@

    We resort to \textsc{CT-Form} on the right.
    This is well-formed,
    since $\Tranty(\nonterm G \vdash \nonterm{ts}_1)$ produces a derivation of $\nonterm{GC} \vdash \nonterm{tc}_1 : \star$,
    and similarly we get a derivation of $\nonterm{GC}, \gterm x : \nonterm{tc}_1 \vdash \nonterm{tc}_2 : \star$ from
    the other invocation of $\Tranty$.
  \item \textsc{TWF-ADT}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\\\
                   \overline{\nonterm G \vdash \nonterm{ts}_i}^i
                  }
                  {\nonterm G \vdash \langle \overline{l_i : \nonterm{ts}_i}^i \rangle}
          \longmapsto
        \inferrule{\overline{\Tranty(\nonterm G \vdash \nonterm{ts}_i)}^i}
                  {\nonterm{GC} \vdash \langle \overline{l_i : \nonterm{tc}_i}^i \rangle : \star}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok),                                         \\
                   &\nonterm{tc}_i = \tranty(\nonterm G \vdash \nonterm{ts}_i).
    \end{align*}

@@

    To show that the right-hand side is well-defined,
    we need to show that $x : \nonterm{tc} \in \nonterm{GC}$ holds.

>>>>

The syntax of the core language is presented in \figref{core_syntax},
the core $\lambda C$ typing rules are defined in \figref{core_typing}
and the rules for our extensions are defined in \figref{core_typing_exts}.
Note that the \textsc{CT-ADTCase} rule uses $\equiv$ denoting the equality
which is a derived form defined later.

@@

Since $\Tranctx$ is defined on the derivations of $\nonterm G \ctxok$,
consider the last rule used in a derivation:
\begin{itemize}
  \item \textsc{TCTX-Empty}.
    \begin{align*}
      & \inferrule{~}
                  {\emptyset \ctxok}
          \longmapsto
        \emptyset.
    \end{align*}
    The base case is trivial: the empty context is mapped to the empty context.
  \item \textsc{TCTX-Bind}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\ \nonterm G \vdash \nonterm{ts}}
                  {\nonterm G , \gterm x : \nonterm{ts} \ctxok}
          \longmapsto
        \Tranctx(\nonterm G \ctxok), x : \tranty(\nonterm G \vdash \nonterm{ts}).
    \end{align*}

@@

More formally, consider the last rule used in the derivation of $\nonterm G \vdash \nonterm{ts}$:
\begin{itemize}
  \item \textsc{TWF-TrueRef}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok}
                  {\nonterm G \vdash \{ v : B | \top \}}
          \longmapsto
        \inferrule{\cdots}
                  {\nonterm{GC} \vdash (\hat B, \unitc \equiv \unitc) : \star}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok).
    \end{align*}

    By \cref{lma:non_dep_pair_typing}
    $(\hat B, \unitc \equiv \unitc)$ has type $\star$ in $\nonterm{GC}$
    if $\hat B$ and $\unitc \equiv \unitc$ are well-typed in $\nonterm{GC}$,
    which, in turn, is easily checked.
    Thus, there exists a derivation of $\nonterm{GC} \vdash (\hat B, \unitc \equiv \unitc) : \star$,
    and it is effectively constructible, although omitted for brevity.
  \item \textsc{TWF-Base}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\\\
                   \nonterm G ' \vdash \nonterm{es}_1 : \{ v_2 : B' | \nonterm{r}_1 \} \\\\
                   \nonterm G ' \vdash \nonterm{es}_2 : \{ v_2 : B' | \nonterm{r}_2 \} \\\\
                   \gterm x \fresh}
                  {\nonterm G \vdash \{ v : B | \nonterm{es}_1 = \nonterm{es}_2 \}}
          \longmapsto
        \inferrule{\Tranterm(\nonterm G ' \vdash \nonterm{es}_1 : \{ v_2 : B' | \nonterm{r}_1 \}) \\\\
                   \Tranterm(\nonterm G ' \vdash \nonterm{es}_2 : \{ v_2 : B' | \nonterm{r}_2 \}) \\\\
                   \gterm x \fresh \\\\
                   \cdots}
                  {\nonterm{GC} \vdash (v : \hat B \dast \lambda v : \hat B. \nonterm{ec}_1 \equiv \nonterm{ec}_2) : \star}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok), \\
                   &\nonterm G ' = \nonterm G , v : \{ v_1 : B | \top \}, \\
                   &\nonterm{ec}_i =
                     \tranterm(\nonterm G ' \vdash \nonterm{es}_i : \{ v_2 : B' | \nonterm{r}_i \}).
    \end{align*}

@@

    The rule \textsc{CT-Form} is used on the right.
    This is admissible,
    since $\Tranty(\nonterm G \vdash \nonterm{ts}_1)$ produces a derivation of $\nonterm{GC} \vdash \nonterm{tc}_1 : \star$,
    and similarly there is a derivation of $\nonterm{GC}, \gterm x : \nonterm{tc}_1 \vdash \nonterm{tc}_2 : \star$
    resulting from the other invocation of $\Tranty$.
  \item \textsc{TWF-ADT}.
    \begin{align*}
      & \inferrule{\nonterm G \ctxok \\\\
                   \overline{\nonterm G \vdash \nonterm{ts}_i}^i
                  }
                  {\nonterm G \vdash \langle \overline{l_i : \nonterm{ts}_i}^i \rangle}
          \longmapsto
        \inferrule{\overline{\Tranty(\nonterm G \vdash \nonterm{ts}_i)}^i}
                  {\nonterm{GC} \vdash \langle \overline{l_i : \nonterm{tc}_i}^i \rangle : \star}, \\
      \text{where}~&\nonterm{GC} = \Tranctx(\nonterm G \ctxok),                                         \\
                   &\nonterm{tc}_i = \tranty(\nonterm G \vdash \nonterm{ts}_i).
    \end{align*}

@@

    To show that the right-hand side is well-defined,
    it suffices to show that $x : \nonterm{tc} \in \nonterm{GC}$ holds.
